---
title: "Problem Set 2"
subtitle: "Mini Data Science"
author: "Nadya Paschall"
institute: "Vanderbilt University"
date: "Due Date: 2024-09-06"
output:
  html_document: default
---

```{r,include=F}
knitr::opts_chunk$set(error=TRUE)
```


## Getting Set Up

Open `RStudio` and create a new RMarkDown file (`.Rmd`) by going to `File -> New File -> R Markdown...`.
Accept defaults and save this file as `[LAST NAME]_ps2.Rmd` to your `code` folder.

Copy and paste the contents of this `.Rmd` file into your `[LAST NAME]_ps2.Rmd` file. Then change the `author: [Your Name]` to your name.

We will be using the `sc_debt.Rds` file from the course [github page](https://github.com/jbisbee1/DS1000_F2024/blob/main/data/sc_debt.Rds).

All of the following questions should be answered in this `.Rmd` file. There are code chunks with incomplete code that need to be filled in. 

This problem set is worth 5 total points, plus 1 extra credit point. The point values for each question are indicated in brackets below. To receive full credit, you must have the correct code. In addition, some questions ask you to provide a written response in addition to the code.

You are free to rely on whatever resources you need to complete this problem set, including lecture notes, lecture presentations, Google, your classmates...you name it. However, the final submission must be complete by you. There are no group assignments. To submit, compiled the completed problem set and upload the PDF file to Brightspace on Friday by midnight. Instructions for how to compile the output as a PDF can be found in [Problem Set 0](https://github.com/jbisbee1/DS1000_F2024/blob/main/Psets/ds1000_pset_0.pdf) and in this [gif tutorial](https://github.com/jbisbee1/DS1000_F2024/blob/main/Psets/save_as_pdf.gif). 

*Note that the TAs and professors will not respond to Campuswire posts after 5PM on Friday, so don't wait until the last minute to get started!*

**Good luck!**

*Copy the link to ChatGPT you used here: _________________



## Question 0 [0 points]
*Require `tidyverse` and load the `sc_debt.Rds` data by assigning it to an object named `df`.*
```{r}
require(tidyverse) # Load tidyverse
df <- read_rds('https://github.com/jbisbee1/DS1000_F2024/raw/main/data/sc_debt.Rds') # Load the dataset
```


## Question 1 [1 point]
*Research Question: Do students who graduate from smaller schools (i.e., schools with smaller student bodies) make more money in their future careers? Before looking at the data, write out what you think the answer is, and explain why you think so.*

>- 
#I do not think that students who go to smaller schools make more. I am thinking from the business perspective and going with if you have an approprite major or qualifications before checking what school it is. 

## Question 2 [1 point]
*Based on this research question, what is the outcome / dependent / $Y$ variable and what is the explanatory / independent / $X$ variable? What are their average values in the data?*
```{r}
df %>%
  select(md_earn_wne_p6)
```

```{r}
df %>%
  summarise(avg_ugds = mean(ugds, na.rm=TRUE))
```
```{r}
df %>%
  summarise(avg_earn = mean(md_earn_wne_p6, na.rm = TRUE))
```

>- Write answer here

#The X variable in this case would be the number of students at the school. The Y would be the median earnings of undergraduates. The average value for earnings is $33,028 and 4,861 undergraduate size.

## Question 3 [1 point]
*Create the scatterplot of the data that analyzes your hypothesis, along with a line of best fit. Then, describe the result. Is your answer to the research question supported?*
```{r}
df %>%
  ggplot(aes(x = ugds, # Put the explanatory variable on the x-axis
             y = md_earn_wne_p6,)) + # Put the outcome variable on the y-axis
  geom_point(color = "blue") + # Create a scatterplot
  geom_smooth(method = "lm",
             color = "purple", linetype = "dashed")+ # Add line of best fit
  labs(title = 'Scatterplot on School Population and Earnings', 
       x = 'Number of Undergrads',
       y = 'Median Earnings of Undergrads')
```

>- Write answer here. 
#Based off my graph, there seems to be a cluster with the number of undergraduates being less that 12,000 students and also earnings being around $30,000.As for supporting my original answer, I would say this does. There actually seems to be a trend toward the median population size as compared to a higher or lower population thus proving the research question wrong but not necessarily me right either since I guess a graph of my prediction would be scatter and more disorganized than this graph which does seem to have a trend.

## Question 4 [1 point]
*Does this relationship change by whether the school is a research university? Using the filter() function, create two versions of the plot, one for research universities and the other for non-research universities. What do you find?*

```{r}
df %>%
  filter( research_u == 0) %>% # Filter to non-research universities
  ggplot(aes(x = ugds, # Put the explanatory variable on the x-axis
             y = md_earn_wne_p6)) +  # Put the outcome variable on the y-axis
  geom_point(color = "green") + # Create a scatterplot
  geom_smooth(method = "lm", color = "yellow", linetype = "dashed") + # Add line of best fit
  labs(title = 'Scatterplot on Non-research School Population and Earnings',
       x = 'Number of Undergrads',
       y = 'Median Earnings of Undergrads')

df %>%
  filter( research_u == 1) %>% # Filter to research universities
  ggplot(aes(x = ugds, # Put the explanatory variable on the x-axis
             y = md_earn_wne_p6)) +  # Put the outcome variable on the y-axis
  geom_point(color = "blue") + # Create a scatterplot
  geom_smooth(method = "lm", color = "pink", linetype = "solid") + # Add line of best fit
  labs(title = 'Scatterplot on Research School Population and Earnings', 
       subtitle = 'In the United States', 
       x = 'Number of Undergrads',
       y = 'Median Earnings of Undergrads')
```

>- Write answer here
#For non-research schools, the graph looks almost identical to the graph displaying all schools however the line of best fits slightly lowers. This must mean that the opposite happens for the research school. This graphs limit is at 30,000 undergrads because the school populations does not go past 40,000 unlike non-research schools who goes up to almost 100,000 and other high populations. With the research schools, a proportion of them are above that $50,000 earnings but majority are scattered around the same fraom of $30,000 to $45,000. The samel can also be said about the non-research schools. There is that group that earn more than the other 80% but I wouldn't say graphically that it shows an advantage to smaller schools because those high schools are reaching that same median income.

## Question 5 [1 point]
*Instead of creating two separate plots, color the points by whether the school is a research university. To do this, you first need to modify the research_u variable to be categorical (it is currently stored as numeric). To do this, use the mutate command with `ifelse()` to create a new variable called `research_u_cat` which is either "Research" if `research_u` is equal to 1, and "Non-Research" otherwise.*
```{r}
df <- df %>%
  mutate(research_u_cat = ifelse(research_u == 1, # Logic
                                 "Research", # Value if true
                                 "Non-research"))# Value if false

df %>%
  ggplot(aes(x = ugds, # Put the explanatory variable on the x-axis
             y = md_earn_wne_p6,      # Put the outcome variable on the y-axis
             color = research_u_cat)) + # Color by whether it is a research university
  geom_point() + # Create a scatterplot
  geom_smooth(method = "lm", color = "red", linetype = "solid") + # Add line of best fit
  labs(title = 'Nonresearch vs Research School Population and Earnings', # give the plot meaningful labels to help the viewer understand it
       subtitle = 'In the United States', 
       x = 'Number of Undergrads',
       y = 'Median Earnings of Undergrads')
```
# Fix this 

## Extra Credit [2 points]
*Write a short paragraph discussing your findings. What do you think is going on in these data?*

>- Write answer here

#Overall, I do not think either graphs completely prove the research question. I believe that there is no advantage to the less student body as compared to the higher populations. Majority of the schools gather around the $30,000 to $45,000. Thoses that are above that limit ranges from close to 0 students to almost 70,000. This thus disproves the origianl question. As for research vs nonresearch, there is a trend  of the research schools being slightly above that line of best fit. So I suggest that maybe it is research schools that allow for more earnings rather than population size.
